ChemKED: a human- and machine-readable data standard for chemical kinetics experiments

### 2017-09-29 rebuttal.tex

    However, I suggest the publication of an improved and shortened version of the manuscript in
    IJCK. The reason is that a permanent and laboratory-independent storage of experimental data is
    an important issue and I agree that a public discussion promotes the acceptance of a widely used
    common format.
    %
\end{reviewer}

\begin{response}
    %
    We would like to thank the reviewer for their comments. Detailed responses to each comment are
    below.
    %
\end{response}

\begin{reviewer}
    %
    I think that the paper contains too much information science details. Still, as the authors
    note, the description is not comprehensive and they refer to the user manual. I suggest the
    creation of a significantly abridged version that contains the basic information technology
    points only and does not discuss the details. For example, Chapter 4 could be deleted and
    Chapters 2 and 3 may also be shortened.
    %
\end{reviewer}

\begin{response}
    %
    Our intention is to describe the system to an audience that may not be familiar with the
    technical details of how the system is implemented. Thus, we feel that the length of the
    manuscript is a necessary outcome of providing the background for this information. However, the
    reviewer is correct that it is not ideal to leave much of the information in the documentation
    only. In combination with the suggestions of Reviewer 2, we have expanded the paper to provide
    more detail where necessary.
    %
\end{response}

@@

\begin{response}

\end{response}

\begin{reviewer}
    %
    13.~Page 6, line 12 describes the initial efforts of building an open repository of ChemKED
    files, on GitHub. Is this the most efficient way for making a publically available database? One
    issue with this is the fact that adding files is done via pull requests, which have to be
    verified and approved by the owners of the repository. Therefore a certain delay is inevitable.
    %
\end{reviewer}

\begin{response}
    We agree that submitting files to the ChemKED database on GitHub will inevitably incur some delay,
    however
\end{response}

### 2017-09-29 rebuttal.tex

\begin{response}
    We attempted to access the GitLab repository at \url{https://git.rwth-aachen.de/ITV/FlameMaster}, but
    it is a closed repository. Thus, while it may be freely available to those with access, it is not
    \emph{openly} available.
\end{response}

\begin{reviewer}
    %
    p.~2 ``Second, the XML format is intended to be a machine-readable markup language rather than a
    data format, and its lack of human readibility presents a barrier to creating and working with
    database files.''

    XML is a text format and it is fully human readable, similarly to YMAL. Even a
    first-time user is able to read the chemical content of a  PrIMe XML file and an experienced
    user is able to write such a file.
    %
\end{reviewer}

\begin{response}
    %
    We certainly recognize that XML is typically written in ASCII or UTF-8 (plain text) files, which
    a human can hypothetically decipher. Our argument is that XML is significantly harder for a
    human to read and write than YAML. This is primarily because XML has a lower signal-to-noise
    ratio than YAML; it requires many more characters to express a data structure in XML than in
    YAML. For instance, XML requires opening and closing tags, with the associated angle brackets,
    for every field while YAML does not; in XML, information can be specified as properties of tags
    or as elements (requiring the user to know where to look for information) while YAML has only 3 basic data structures (keys, values, and lists).

    We note that although this verbosity in XML is a disadvantage when humans attempt to read or
    write such files, it allows much greater flexibility in specifying how a document should be
    marked up and provides XML with many advantages in other areas. However, since YAML is intended
    as a simple data format, and has a much simpler overall format, we argue that it is therefore
    much easier for humans to work with.

    We have modified the text to clarify this point, including describing the lack of human
    readability as a formatting issue (but one that is inherent to XML), rather than a technical
    issue.
    %
\end{response}

### 2017-09-29 pyked.tex
Work on #7, #23. Clarify that the problem with XML is verbosity

PrIMe~\autocite{Frenklach:2007bm,You:2011hy} has a number of features that make
it a strong standardization format, but it suffers from several flaws that
prevent wide adoption. First, the PrIMe standard does not require or support all
the information needed to simulate an experiment, including a machine-readable
definition of ignition or a standard way to express detailed facility-specific
effects necessary to properly simulate certain experiments. Second, the XML
format is intended to be a machine-readable markup language rather than a data format, and
its lack of human readibility presents a barrier to creating and working with
database files. Furthermore, the
focus on using internal identifiers for bibliographic references and
species\slash reactions, rather than standard identifiers (e.g., DOIs for
scholarly products, InCHI or SMILES for species), limits usability outside
the PrIMe ecosystem. Finally, the closed and opaque natures of the PrIMe standard
and associated Data Warehouse limit contributions and community development of
the PrIMe standard and compatible data files and tools.

More recently, Varga and coworkers~\autocite{Varga2015a,Varga2017} developed the
ReSpecTh standard, which builds on the PrIMe format. ReSpecTh adds important
features that make files better standalone representations of experimental
data---i.e., more informative by themselves rather than in concert with a larger
system. For example, where PrIMe uses internal bibliographic references,
ReSpecTh adds a field for typical bibliographic data. The ReSpecTh standard also
provides machine-readable formats for describing ignition experiments, including
a field for the definition of autoignition and the ability to specify
facility-specific effects. However, ReSpecTh experimental files are another
XML-based format, and, as such, suffer from the same usability issues as PrIMe.
Moreover, the standard remains inaccessible to the public, requiring potential users
to register with the standard's authors using a valid institutional email address
to access the specification and the growing database.

>>>>

PrIMe~\autocite{Frenklach:2007bm,You:2011hy} has a number of features that make
it a strong standardization format, but it suffers from several flaws that
prevent wide adoption. First, the PrIMe standard does not require or support all
the information needed to simulate an experiment, including a machine-readable
definition of ignition or a standard way to express detailed facility-specific
effects necessary to properly simulate certain experiments. Second, the XML
format is intended to be a machine-readable markup language rather than a data format.
As a general-purpose document markup language, XML is able to represent very detailed
information about a dataset, but implementing this detail requires XML-formatted
files to be significantly more verbose than files formatted with a language focused
only on representing data structures. The verbosity inherent in the XML format
limits the human readability of database files constructed in XML and presents a barrier to
creating and working with them. Furthermore, PrIMe uses
internal identifiers for bibliographic references and
species\slash reactions, rather than standard identifiers (e.g., DOIs for
scholarly products, InCHI or SMILES for species). This limits the usability of PrIMe files outside
the PrIMe ecosystem. Finally, the closed and opaque natures of the PrIMe standard
and associated Data Warehouse limit contributions and community development of
the PrIMe standard and compatible data files and tools.

More recently, Varga and coworkers~\autocite{Varga2015a,Varga2017} developed the
ReSpecTh standard, which builds on the PrIMe format. ReSpecTh adds important
features that make files better standalone representations of experimental
data---i.e., more informative by themselves rather than in concert with a larger
system. For example, where PrIMe uses internal bibliographic references,
ReSpecTh adds a field for typical bibliographic data. The ReSpecTh standard also
provides machine-readable formats for describing ignition experiments, including
a field for the definition of autoignition and the ability to specify
facility-specific effects. It can also describe speciation measurements in
perfectly stirred reactors, flow reactors, and burner-stabilized flames, as well as
burning velocity data. However, ReSpecTh experimental files are another
XML-based format, and, as such, suffer from the same usability issues as PrIMe.
Moreover, the standard remains inaccessible to the public, requiring potential users
to register with the standard's authors using a valid institutional email address
to access the specification and the growing database.
